openTSNE: a modular \proglang{Python} library for t-SNE dimensionality reduction and embedding

### 2022-03-28 paper/main.tex
Add iris tutorial

\subsection{Example}
\subsubsection{Installation}

@@

To verify that that \opentsne\ was installed correctly, try to import the module:
\begin{CodeChunk}
\begin{CodeInput}
import openTSNE
print(openTSNE.__version__)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
from openTSNE import TSNE

x: np.ndarray = ...

# Vanilla t-SNE
tsne_standard = TSNE().fit(x)

# Adapt parameters to better highlight small clusters in the data
tsne_extensions = TSNE(perplexity=50, exaggeration=2, dof=0.8).fit(x)

# Use multiple perplexity values for better local/global structure trade-off
tsne_multiscale = TSNE(perplexity=[50, 500]).fit(x)

# Use the generated embeddings to add new data points
new_embedding = tsne_extensions.transform(x_new)
\end{CodeInput}
\end{CodeChunk}
The code snippet considers a \pkg{numpy} array or \pkg{scipy} sparse matrix \texttt{x} and creates three \texttt{TSNEEmbedding} objects. These represent fully-optimized t-SNE embeddings, and \opentsne\ internally stores them as \pkg{numpy} arrays. These can be manipulated and visualized using standard Python tools. We then select one of these embeddings and use it to embed new data points from \texttt{x\_new} into the embedding space while keeping the reference embedding fixed. Additionally, \opentsne\ provides a more advanced interface that enables the user to change every aspect of the t-SNE optimization procedure. We describe this interface in detail in the \opentsne's on-line documentation.

>>>>

\subsection{Example}

\subsubsection{Installation}

@@

To verify that that \opentsne\ was installed correctly, try to import the module:
\begin{CodeChunk}
\begin{CodeInput}
>>> import openTSNE
>>> print(openTSNE.__version__)
0.6.2
\end{CodeInput}
\end{CodeChunk}

@@

We demonstrate how to use \opentsne\ on the Iris data set from \citet{anderson1936species}.
This small data set describes 150 iris flowers containing four variables (sepal
length, sepal width, petal length, petal width) and their species (Iris Setosa,
Iris Versicolor, Iris Virginica), which is the target variable.
This data set is included in \pkg{scikit-learn}, making it easy to load:
\begin{CodeChunk}
\begin{CodeInput}
>>> from sklearn import datasets
>>>
>>> iris = datasets.load_iris()
>>> iris.data[:3]
array([[5.1, 3.5, 1.4, 0.2],
       [4.9, 3. , 1.4, 0.2],
       [4.7, 3.2, 1.3, 0.2]])
\end{CodeInput}
\end{CodeChunk}
\pkg{scikit-learn} loads the data set into a \pkg{numpy} array, which can directly
be used by \opentsne:
\begin{CodeChunk}
\begin{CodeInput}
>>> import openTSNE
>>>
>>> embedding = openTSNE.TSNE().fit(iris.data)
>>> embedding[:3]
TSNEEmbedding([[-0.55581917, 17.56162344],
               [-1.76742536, 20.07763453],
               [-0.58444608, 20.00052495]])
\end{CodeInput}
\end{CodeChunk}
\opentsne\ returns an \code{TSNEEmbedding} object, which is a subclass of
\code{np.ndarray}. This allows users to easily manipulate the resulting embedding.
The additional functionality provided by the \code{TSNEEmbedding} object allows
the embedding to be further optimized using different parameter values, or be
used as a reference when obtaining new data points.
For users who do not need this additional functionality, this object can be
converted to a standard \code{np.ndarray} with:
\begin{CodeChunk}
\begin{CodeInput}
>>> import numpy as np
>>>
>>> X_tsne = embedding.view(np.ndarray)
>>> X_tsne[:3]
array([[-0.55581917, 17.56162344],
       [-1.76742536, 20.07763453],
       [-0.58444608, 20.00052495]])
\end{CodeInput}
\end{CodeChunk}
Fig.~\ref{fig:iris} shows a visualization of the resulting t-SNE visualization.
As expected, t-SNE is correctly able to group together samples from the same
species. Additionally, the t-SNE embedding reveals that Iris Setosa is markedly
different from the other two species, while flowers from Iris Versicolor and
Iris Virginica may appear more similar to one another.
\begin{figure*}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{iris}
  \caption{\label{fig:iris}
  Visualization of the obtained t-SNE embedding on the Iris data set. The colors
  correspond to different species of the Iris flower.
}
\end{figure*}
